Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-03-21 manuscript/manuscript.Rmd
added paragraph on Verity's project and extra sentence about p-original

::: {.abstract data-latex="" lang="en"}
When a replication fails, scientists have to decide whether to make a second attempt or move on. Psychology researchers who attempt to replicate studies before building on them often face this decision, given the relatively low rates of replication success in psychology. Here, we report on 17 re-replications of experiments with failed replications reported in Boyce et al. (2023). In 5/17 of these "rescue" projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few replications were rescued with larger sample sizes and minor methodological changes, most re-replications failed again. In many situations, it may be most efficient to stop pursuing an effect after a single failed replication.
:::
TODO FIX CITATION / BIB INFO (re OSC for RP:P but only after we unlink from zotero)

@@

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that one study's population effect comes from the same distribution of population effects as another study or studies [@mathurNewStatisticalMetrics2020a] If we assume no heterogeneity between studies, then this is the same as asking how consistent one study's estimate of the population-level effect is with another study's estimate of the population-level effect. However, meta-analyses of closely related studies, such as those from different sites of multi-site replications, indicate that there is often heterogeneity between studies. We cannot accurately estimate the level of heterogeneity for an effect meta-analytically as we have only a small number of replications per study. Instead, we impute a heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

(TODO other successes or failures we want to discuss? ) (Ben: Emily's project was interesting: the rescue succeeded but with big stimulus-level random effects. Verity's project was also interesting: the previous replication, as well as another in the literature, showed similar but smaller findings and it really looked like she could get the effect with a bigger sample, but she ended up finding totally different results.)

>>>>

::: {.abstract data-latex="" lang="en"}
When a replication fails, scientists have to decide whether to make a second attempt or move on.
Psychology researchers who attempt to replicate studies before building on them often face this decision, given the relatively low rates of replication success in psychology.
Here, we report on 17 re-replications of experiments with failed replications reported in Boyce et al. (2023).
In 5/17 of these "rescue" projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes.
While a few replications were rescued with larger sample sizes and minor methodological changes, most re-replications failed again.
In many situations, it may be most efficient to stop pursuing an effect after a single failed replication.
:::

TODO FIX CITATION / BIB INFO (re OSC for RP:P but only after we unlink from zotero)

@@

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that one study's population effect comes from the same distribution of population effects as another study or studies [@mathurNewStatisticalMetrics2020a] If we assume no heterogeneity between studies, then this is the same as asking how consistent one study's estimate of the population-level effect is with another study's estimate of the population-level effect.
However, meta-analyses of closely related studies, such as those from different sites of multi-site replications, indicate that there is often heterogeneity between studies.
$p$-original allows for such heterogeneity, as governed by a parameter $\tau$.
We cannot accurately estimate the level of heterogeneity for an effect meta-analytically as we have only a small number of replications per study.
Instead, we impute a heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

\@child2018 studied reading times for passages of text with different emotional valences written from either a third-person or second-person perspective.
They found a significant interaction between perspective and valence: participants read positively-valenced passages faster than negatively-valenced passages, but only when they were written from a second-person perspective.
The first replication used a similar sample size to the original study (40) and found an effect that trended in the same direction, but was substantially smaller.
The rescuer identified several possible reasons for this failure, foremost among them that the replication was conducted online while the original was conducted in person.
It is reasonable to expect online studies of reading time to be noisier than in-person studies because the environments participants completed the study in varied much more online.
The rescue attempt was still online, but used a much larger sample size to account for noisier data.
Still, the rescue found a non-significant effect that trended in the opposite direction from the original and first replication.
Even though the first replication trended in the same direction and there was an identifiable reason for the smaller effect size, increasing sample size did not rescue the effect.
